PLAE web app enables powerful searching and multiple visualizations across one million unified single-cell ocular transcriptomes

### 2022-08-31 manuscript_00.Rmd
tighten up heatmap

## Leveraging the high diversity of studies to identify thousands of consistently differentially expressed genes across ocular cell types

@@

Across the cell types of the eye, we find `r diff_tab %>% nrow()` gene - celltype combinations that are significantly differentially expressed. We define significant as a log2 fold change greater than two as well as a p-value (corrected) less than $1 x 10^-5$. We find a mean of 
`r diff_tab_sup$"Homo sapiens" %>% mean() %>% round()`, 
`r diff_tab_sup$"Mus musculus" %>% mean() %>% round()`, and 
`r diff_tab_sup$"Macaca fascicularis" %>% mean() %>% round()` differentially expressed genes in human, macaque, and mouse respectively (`r supTab_cap(name = 'diff_tab_sup_tab', display = 'cite')`). These variable results directly reflect the differing number of independent studies - mouse and human have substantially more than macaque (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). The significant filtered differential table is made available as supplementary file "differential_expression_table.csv.gz" and the full differential tables are available for download at plae.nei.nih.gov -> Data.

@@

To visualize these top markers we select up to eight genes (ordered by the mean log2 fold change across organisms) per well supported cell type for each and plot the log2 fold change of the gene expression relative to all other cell types (`r fig_cap(name = 'consist_genes_heatmap', display = 'cite')`). We see that human, mouse, and macaque have similar expression patterning of the proposed markers across the well supported cell types. The cell types are arranged by the columns by how related the expression patterns are. We see how the differentiated cell types of the retina are grouped together with the exception of the Müller glia, which are next to the RPC from which they derive from. The photoreceptors are also next to each other

>>>>

<!-- ## Leveraging the high diversity of studies to identify thousands of consistently differentially expressed genes across ocular cell types -->

<!-- Across the cell types of the eye, we find `r diff_tab %>% nrow()` gene - celltype combinations that are significantly differentially expressed. We define significant as a log2 fold change greater than two as well as a p-value (corrected) less than $1 x 10^-5$. We find a mean of  -->
<!-- `r diff_tab_sup$"Homo sapiens" %>% mean() %>% round()`,  -->
<!-- `r diff_tab_sup$"Mus musculus" %>% mean() %>% round()`, and  -->
<!-- `r diff_tab_sup$"Macaca fascicularis" %>% mean() %>% round()` differentially expressed genes in human, macaque, and mouse respectively (`r supTab_cap(name = 'diff_tab_sup_tab', display = 'cite')`). These variable results directly reflect the differing number of independent studies - mouse and human have substantially more than macaque (`r supTab_cap(name = 'well_supported_celltypes_tab', display = 'cite')`). The significant filtered differential table is made available as supplementary file "differential_expression_table.csv.gz" and the full differential tables are available for download at plae.nei.nih.gov -> Data.  -->

@@

@@

To visualize these top markers we select up to eight genes (ordered by the mean log2 fold change across organisms) per well supported cell type for each and plot the log2 fold change of the gene expression relative to all other cell types (`r fig_cap(name = 'consist_genes_heatmap', display = 'cite')`). We see that human, mouse, and macaque have similar expression patterning of the proposed markers across the well supported cell types. The cell types are arranged by the columns by how related the expression patterns are. We see how the differentiated cell types of the retina are grouped together with the exception of the Müller glia, which are next to the RPC from which they derive from. The photoreceptors are also next to each other.

Aasdfasdfasdfadsf
